---
title: "615 Final Project"
author: "Haoran Cui"
format: pptx
editor: visual
---

```{r}
library(tidyverse)
library(shiny)
library(leaflet)
```

## Loading and Cleaning Data

This code processes a dataset by cleaning column names to retain only the year, filtering for specific indicators like GDP and population, and reshaping the data into a tidy format for analysis. It extracts data for the years 2000 to 2023, ensures columns like "Year" and "Value" are numeric, and removes rows with missing values. The result is a clean, long-format dataset suitable for visualizations or further analysis.

```{r}
data <- read.csv("1.csv")

# Clean column names to extract only the year
colnames(data) <- gsub("^X(\\d{4})..YR\\1\\.$", "\\1", colnames(data))

# Define the indicators of interest
selected_indicators <- c('GDP (current US$)', 'GDP growth (annual %)', 
                         'GDP per capita (current US$)', 'Population, total', 
                         'Population growth (annual %)', 'Life expectancy at birth, total (years)')

# Filter and reshape the data
filtered_data <- data %>%
  filter(Series.Name %in% selected_indicators) %>%
  select(Series.Name, any_of(as.character(2000:2023))) %>%
  pivot_longer(cols = any_of(as.character(2000:2023)), names_to = "Year", values_to = "Value") %>%
  mutate(
    Year = as.numeric(Year),  # Ensure Year is numeric
    Value = as.numeric(gsub("[^0-9.-]", "", Value))  # Clean Value column
  ) %>%
  drop_na(Value)
```

## **Overview of the App's Purpose**

The app organizes information about Aruba into multiple sections (tabs) to provide a structured, user-friendly way of understanding its economic, demographic, geographic, and ecological characteristics. Users can interact with the app to explore data and gain insights.

## **What Each Section Does**

1.  **Introduction Tab**:

    -   Provides a general overview of Aruba.

    -   Includes text about Aruba's key features, such as its geography, economy, and government.

    -   Displays an image of Aruba to visually complement the description.

2.  **Biodiversity Tab**:

    -   Highlights Aruba's unique ecosystems, including its flora and fauna.

    -   Explains features like desert landscapes and coral reefs, along with native species.

    -   Includes two images to visually showcase Aruba's natural beauty.

3.  **Interactive Map Tab**:

    -   Offers an interactive map of Aruba.

    -   Users can zoom and pan the map to explore the island.

    -   Displays a marker for Oranjestad, Aruba's capital, to provide geographic context.

4.  **Economic Indicators Tab**:

    -   Allows users to explore Aruba's economic performance.

    -   Users can select specific economic indicators like GDP or GDP growth.

    -   Displays a time-series graph that shows how the selected indicator has changed over the years.

5.  **Population Statistics Tab**:

    -   Focuses on demographic trends in Aruba.

    -   Users can choose variables like total population, population growth, or life expectancy.

    -   Displays a graph that tracks the chosen variable over time, helping to visualize changes.

6.  **Comparison Tab**:

    -   Compares Aruba with two other islands: Dominica and Jamaica.

    -   Text highlights key differences and similarities, such as GDP, population, economic structure, and vulnerabilities.

    -   Includes relevant images to make the comparison more engaging.

7.  **SWOT Analysis Tab**:

    -   Provides a structured analysis of Aruba's:

        -   **Strengths** (e.g., strong tourism sector, high life expectancy).

        -   **Weaknesses** (e.g., small population, reliance on tourism).

        -   **Opportunities** (e.g., potential for economic diversification).

        -   **Threats** (e.g., climate change, natural disasters).

    -   Summarizes insights into Aruba’s position and challenges in an easily digestible format.

## **What Makes It Interactive**

The app allows users to:

-   Explore data dynamically by selecting variables (e.g., GDP, population growth).

-   Interact with an online map to see Aruba's geographic layout.

-   Read comparative analyses and view related images for a better contextual understanding.

```{r}
# Add resource path for images
addResourcePath("images", "C:/Users/cuih1/OneDrive/Desktop/615/Aruba data")

# UI
ui <- navbarPage(
  "Aruba Analysis",
  tabPanel("Introduction",
           fluidPage(
             titlePanel("About Aruba"),
             fluidRow(
               column(6, 
                      p("Aruba is a Caribbean island located off the coast of Venezuela, known for its white-sand beaches, turquoise waters, and year-round sunny climate. Aruba is part of the Kingdom of the Netherlands, with its own government and currency. Its economy heavily relies on tourism, aloe production, and oil refining."),
                      p("Capital City: Oranjestad"),
                      p("Currency: Aruban Florin (AWG)"),
                      p("Land area: 180 square kilometers")
               ),
               column(6, img(src = "images/Aruba_1.png", width = "100%"))
             )
           )
  ),
  tabPanel("Biodiversity",
           fluidPage(
             titlePanel("Aruba's Biodiversity"),
             p("Aruba's natural ecosystems include desert landscapes, coral reefs, and unique flora like the Divi-Divi tree. Native species include sea turtles, lizards, and exotic birds."),
             fluidRow(
               column(6, img(src = "images/Aruba_2.jpg", width = "100%")),
               column(6, img(src = "images/Aruba_3.jpg", width = "100%"))
             )
           )
  ),
  tabPanel("Map",
           fluidPage(
             titlePanel("Interactive Map of Aruba"),
             leafletOutput("arubaMap", height = 500)
           )
  ),
  tabPanel("Economic Indicators",
           fluidPage(
             titlePanel("Economic Indicators"),
             selectInput("econVar", "Choose a variable to display:", 
                         choices = c("GDP (current US$)", "GDP growth (annual %)")),
             plotOutput("econPlot")
           )
  ),
  tabPanel("Population Statistics",
           fluidPage(
             titlePanel("Population Trends"),
             selectInput("popVar", "Choose a variable to display:", 
                         choices = c("Population, total", "Population growth (annual %)", "Life expectancy at birth, total (years)")),
             plotOutput("popPlot")
           )
  ),
  tabPanel("Comparison",
           fluidPage(
             titlePanel("Comparison with Dominica and Jamaica"),
             h4("Aruba vs Dominica"),
             p("- Aruba has a significantly higher GDP compared to Dominica due to its developed tourism infrastructure and stable economy. Dominica, on the other hand, relies more on agriculture and eco-tourism."),
             p("- In terms of population, Dominica is much smaller than Aruba with approximately 70,000 people, while Aruba has around 120,000. Both islands share vulnerabilities to natural disasters such as hurricanes."),
             img(src = "images/Dominica.png", width = "100%"),
             h4("Aruba vs Jamaica"),
             p("- Jamaica, with a population of nearly 3 million, far exceeds Aruba in terms of size and workforce. However, Aruba's GDP per capita is significantly higher due to its higher-income economy."),
             p("- Jamaica's economy is more diversified, including agriculture, mining, and manufacturing, while Aruba remains heavily dependent on tourism."),
             p("- Both Aruba and Jamaica face challenges from climate change, but Jamaica has more resources to invest in long-term solutions given its larger economy."),
             img(src = "images/Jamaica.png", width = "100%")
           )
  ),
  tabPanel("SWOT Analysis",
           fluidPage(
             titlePanel("SWOT Analysis of Aruba"),
             h4("Strengths"),
             p("- GDP: Aruba's current GDP indicates a relatively strong economy supported by tourism and industry."),
             p("- GDP Growth: A positive annual growth rate of 3.5% shows economic resilience."),
             p("- Life Expectancy: At 76.5 years, Aruba has a higher-than-average life expectancy compared to other SIDS (Small Island Developing States)."),
             h4("Weaknesses"),
             p("- Population Growth: A modest growth rate of 1.2% may indicate limited workforce expansion."),
             p("- Population Size: With a small population of 120,000, Aruba faces challenges scaling its economy and workforce for sustainable growth."),
             p("- Dependence on Tourism: The GDP structure heavily relies on external factors like global tourism, making it vulnerable to global economic shocks (e.g., pandemics)."),
             h4("Opportunities"),
             p("- Diversification: Economic diversification into renewable energy and eco-tourism can reduce reliance on traditional tourism."),
             p("- Sustainability: With a stable GDP growth rate, Aruba can invest in green initiatives to align with global environmental trends."),
             p("- Life Expectancy: A higher life expectancy can attract retirees, boosting sectors like healthcare and real estate."),
             h4("Threats"),
             p("- Climate Change: Rising sea levels and hurricanes pose severe threats to the island's economy and infrastructure."),
             p("- Economic Shocks: Tourism dependency makes Aruba susceptible to external shocks like pandemics or global recessions."),
             p("- Resource Scarcity: Aruba’s small land size limits natural resources for self-sustained growth.")
           )
  )
)

# Server
server <- function(input, output, session) {
  # Interactive Map
  output$arubaMap <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      setView(lng = -69.9683, lat = 12.5211, zoom = 10) %>%
      addMarkers(lng = -69.9683, lat = 12.5211, popup = "Aruba: Oranjestad")
  })

  # Economic Indicators Plot
  output$econPlot <- renderPlot({
    req(input$econVar)  # Ensure a variable is selected
    econ_data <- filtered_data %>% filter(`Series.Name` == input$econVar)
    ggplot(econ_data, aes(x = Year, y = Value)) +
      geom_line(color = "blue", size = 1) +
      theme_minimal() +
      labs(
        title = paste("Trends in", input$econVar),
        y = input$econVar,
        x = "Year"
      )
  })

  # Population Statistics Plot
  output$popPlot <- renderPlot({
    req(input$popVar)  # Ensure a variable is selected
    pop_data <- filtered_data %>% filter(`Series.Name` == input$popVar)
    ggplot(pop_data, aes(x = Year, y = Value)) +
      geom_line(color = "red", size = 1) +
      geom_point(color = "red", size = 2) +
      theme_minimal() +
      labs(
        title = paste("Trends in", input$popVar),
        y = input$popVar,
        x = "Year"
      )
  })
}

# Run App
shinyApp(ui, server)

```

## **Why This App is Valuable**

The app serves as a tool to:

-   Educate users about Aruba's unique features and challenges.

-   Provide interactive visualizations that make complex data easier to understand.

-   Support comparative analysis to place Aruba in a broader context alongside other islands.

-   Present a holistic view of Aruba, combining text, visuals, and interactivity for a richer learning experience.

This makes the app useful for education, research, or decision-making related to Aruba's development and sustainability.

## Reference

<https://databank.worldbank.org/reports.aspx?source=2&country=ABW>

<https://www.nytimes.com/2020/03/12/travel/what-to-do-36-hours-in-aruba.html>

<https://www.visitaruba.com/blog/about-aruba/mangroves-and-wildlife-environmental-awareness-from-aruba-pt-ii/>

<https://www.visitaruba.com/blog/things-to-do/arubas-chasing-coral-event-a-campaign-to-save-our-coral-reefs-and-inspire-action/>
